---
title: "01_BRUV_initial_wrangle"
output: html_document
date: '2022-06-22'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This markdown uses the point measurement file associated with each BRUV drop to match observations in Maxn field. For observations that were not in the Eventmneasure species file, often the comment option is used to record the species name quickly. Maxn exported data however, does not contain the associated "comment" and so this must be extracted from he point measurement file.

Best practice to make this wrangling easier in future is to update the species file regularly as you go.

After lots of wrangling, the easiest solution that I could come up with was to manually calculate maxn from the point count data. This ensures that any observation that might match family and genus, but not necessarily species, is assessed based on the frame number and the file name. After this, applying the "comment" species name to "species is simple. Still import the maxn file to double check the same number of observations are returned by the manual maxn calculation

#Housekeeping
```{}
rm(list=ls())
#dev.off()
getwd()
```

#Libraries
```{r}
library(tidyverse)
library(readr)
```


#Import data
Files need to have been saved as csv, all excel modifications made and renamed as per analysis SOP instructions.
```{r,echo=FALSE, results=FALSE, include=FALSE}
# We need the reef, site name as a new variable to identify the source of the data
# Pulls each file in, merges and creates a filename variable. 

path <- "../0_data/BRUV_data"

files <- list.files(path, pattern="csv", full.names=TRUE) %>%
    set_names()

data <- files %>% map_dfr(read_csv, .id="filename")

```


# Tidy variables
```{r}
data <- data %>% mutate(
  filename=str_sub(filename,21,-5)) %>% 
  separate(filename,c("Reef","Site","Drop","Count_type","Survey"), "_", remove = FALSE) %>% 
  mutate(
  Transect=as.factor(Period),
  filename = as.factor(filename),
  Reef = as.factor(Reef),
  Site =  as.factor(paste(Reef, Site, sep = "_")),
  Drop = as.factor(Drop),
  Count_type = as.factor(Count_type),
  Survey=as.factor(case_when(grepl("FEB2021", Survey) ~ "Feb_21",
                             grepl("OCT2021", Survey) ~ "Oct_21")),
  Family = as.factor(Family),
  Genus = as.factor(Genus),
  Species =as.factor(Species))

```

# Change site names in the Site Factor to the R code site names

```{r}
levels(data$Site)

# old_name = new_name

data <- data %>% mutate(Site =recode_factor(Site,'CHILCOTT_CHILCOTT' = 'CHILCOTT_1',
                                           'HERALD_HERALD' = 'HERALD_1',
                                           'HOLMES_HOLMES' = 'HOLMES_1',
                                           'KENN_KENN' = 'KENN_1',
                                           'MARION_MARION' = 'MARION_1',
                                           'WILLIS_WILLIS' = 'WILLIS_1',
                                           "FLINDERS_N" = "FLINDERS_N1",
                                           "LIHOU_N" = "LIHOU_N1",
                                           "LIHOU_S" = "LIHOU_S1",
                                           "SAUMAREZ_N" = "SAUMAREZ_N1",
                                           "WRECK_W" = "WRECK_W1"))
                        
levels(data$Site)

# Ashmore is not just one site - 5 days of drops, 5 different sites - 4/07/22 eventually just manually renamed all the csv files as there was no date ID in the EM files to distinguish. 


```

# De-captialise Reef and Site names
```{r}
data<-data%>% mutate(Site = str_to_title(Site),
                     Reef = str_to_title(Reef))


data$Site<-gsub("(?<=^|_)([a-z])", "\\U\\1", data$Site, perl=TRUE)



```


# Split MAXN and POINT data
Filter using string detect

```{r}
maxn<-data %>% filter(
  str_detect(filename, "MAXN")
)

points<-data %>% filter(
  str_detect(filename, "POINT")
)

```


# Transfer the species name from comments to Species variable in points
```{r}
points<-points %>% 
  mutate(
    Species = case_when(Species == "sp" ~ as.character(Comment),
                        TRUE ~ as.character(Species)),
    Species = as.factor(Species))
```

# Check for remaining NA values in points Species
```{r}
points2<-points %>% filter(is.na(Species)) # 121 Na species
points<-points %>% filter(!is.na(Species)) # 2167 # Non-na species
```

# Addressing missing species issues in points2
```{r}
unique(points2$Comment)

# in Comments:  

# unknown spp is fine

# Aipysurus laevis = olive sea snake - filter observations and save as sep data
# search for "Aipysurus" ---> there are 10 observations in different entry formats (comment, genus and species)

snakes <- data %>% filter(Comment == "Aipysurus laevis" | Genus=="Aipysurus") # note filtered from "data" as 2x observations has "sp" in species so are not detected in the points2 filter
points2<-points2%>%subset(Comment != 'Aipysurus laevis' | is.na(Comment)) # remove snakes from points (2 obs) and points2 (4obs)
points<-points%>%subset(Genus != 'Aipysurus') 
write.csv(snakes, file= "../0_data/BRUV_snake_data.csv", row.names = FALSE)

# Check remaining NA values in points2
points2 %>% filter(is.na(Species)) # 117 obs
points2 %>% distinct(Family, Genus, .keep_all = TRUE) # 38 unique
points3<-points2 %>% distinct(Family, Genus, .keep_all = TRUE)

# All the missing sp look legit, just unknowns, no spelling mistakes etc...

```

# Rebind points and points2
```{r}
# check new points4 matches obs in original points
points4<-rbind(points,points2)
points %>% filter(is.na(Species))
```

# Select frame family genus species from points
```{r}
points2<-points %>% 
  mutate(Species_name=as.factor(paste(Genus,Species, sep = " ")))%>% 
  distinct()

```

# My custom maxn calculation
Picks out distinct observations by frame and chooses the frame with the highest counts. with_ties = FALSE will pick the first frame in a tie.

```{r}
points3<-points2 %>% group_by(filename,Family,Genus, Species) %>%
  slice_max(Number, with_ties = FALSE) %>% 
  mutate(
    MaxN = coalesce(MaxN, Number)
  )

```

At this point check that the number of obs from the manual maxn matches the EM Maxn data.
Note Maxn will also need wrangling to remove seasnakes, missing sp names etc

# Write to csv
```{r}
write.csv(points3, file= "../0_data/BRUV_fish_data.csv", row.names = FALSE)
```



