---
title: "05_biomass_calculations"
author: "Gemma Galbraith"
date: "30/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Used the "Fish_Final" data to calculate the biomass of each fish observation.

#HouseKeeping
```{R}
rm(list=ls())
#dev.off()
getwd()

```

#load libraries
```{r}
library(tidyverse)
library(rfishbase)
```

# Data import
```{r}
fish <-read_csv("../0_data/Fish_Final.csv")

fish <- fish %>% mutate(
  Transect=as.factor(Transect),
  Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Length_mm=as.double(Length_mm),
  Survey=as.factor(Survey),
  Site=as.factor(Site),
  Comment = as.character(Comment),
  Date = as.Date(Date),
  Habitat = as.factor(Habitat),
  Aspect = as.factor(Aspect))

```

# Biomass calcuation for each observation
This takes into account the a and b constants, length in mm and the number of fish/individuals in the observation

W = a x Lb
The units of length and weight in FishBase are centimeter and gram, respectively. Thus when length-weight relationships are not in cm-g, the intercept 'a' is transformed as follows:

a’(cm, g) = a (mm, g)*10^b
a’(cm, g) = a (cm, kg)*1000 
a’(cm, g) = a (mm, mg)*10^b/1000 
a’(cm, g) = a (mm, kg)*10^b*1000

```{r}
# Convert our lengths from mm to cm by dividing by 10
fish<- fish %>% mutate(
  Length_cm= Length_mm/10)

# Column for biomass calculation  W = aL^b
fish<-fish %>% mutate(weight_g = (`Biomass Constant A`*Length_cm^`Biomass Constant B`))

# multiply by number of individuals observed
fish<-fish %>% mutate(Total_Biomass_g = Number*weight_g)

```

# Select variables exculding species names etc and Group by transect
```{r}

biomass.total <- fish %>% select(
Survey,Reef,Site,Transect,T_ID, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity,Total_Biomass_g
)

# Summary per transect total biomas
biomass.total<- biomass.total %>% group_by(Survey,Reef,Site,Transect,T_ID, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity) %>%
  summarise(Total_Biomass_g = sum(Total_Biomass_g))

```


```{r}

write.csv(biomass.total, file= "../0_data/fish_biomass.csv", row.names = FALSE)

```

# Biomass by functional group
```{r}
mermaid.species <-read_csv("../0_data/MERMAID_data/fish-species_GFG.csv")
mermaid.species<- mermaid.species %>% select(Species_name,`Trophic Level`,`Trophic Group`, `Functional Group`)

fish %>% anti_join(mermaid.species, by = "Species_name")

# Figure out which species MERMAID does not have
missing<-fish %>% anti_join(mermaid.species, by = "Species_name")
unique(missing$Species_name)

```

```{r}
# Merge the data
merged.data <- fish %>% 
  left_join(mermaid.species, by ="Species_name")

# Manual application of appropriate trophic groups
merged.data$`Trophic Group`[merged.data$Species_name == "Zebrasoma velifer"] <- ""
merged.data$`Trophic Group`[merged.data$Species_name == ""] <- ""
merged.data$`Trophic Group`[merged.data$Species_name == ""] <- ""



```

# Find NA values
```{r}
summary(is.na(merged.data$`Trophic Group`))

merged.data %>% filter(is.na(`Trophic Group`))

```

# Summarise and Group by trophic group
```{r}
trophic.biomass<-merged.data %>% group_by(Survey,Reef,Site,Transect,T_ID, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity,`Trophic Group`) %>%
  summarise(Total_Biomass_g = sum(Total_Biomass_g))


```
