---
title: "05_biomass_calculations"
author: "Gemma Galbraith"
date: "30/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Uses the "ROV_Fish_Final" data to calculate the biomass of each fish observation.

#HouseKeeping
```{R}
rm(list=ls())
#dev.off()
getwd()

```

#load libraries
```{r}
library(tidyverse)
library(rfishbase)
```

# Data import
```{r}
fish <-read_csv("../2_outputs/ROV_Fish_Final.csv")

fish <- fish %>% mutate(
  Transect=as.factor(Transect),
  Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Length_mm=as.double(Length_mm),
  Survey=as.factor(Survey),
  Site=as.factor(Site),
  Comment = as.character(Comment),
  Date = as.Date(Date),
  Habitat = as.factor(Habitat),
  Aspect = as.factor(Aspect))

```

# Biomass calcuation for each observation
This takes into account the a and b constants, length in mm and the number of fish/individuals in the observation

W = a x Lb
The units of length and weight in FishBase are centimeter and gram, respectively. Thus when length-weight relationships are not in cm-g, the intercept 'a' is transformed as follows:

a’(cm, g) = a (mm, g)*10^b
a’(cm, g) = a (cm, kg)*1000 
a’(cm, g) = a (mm, mg)*10^b/1000 
a’(cm, g) = a (mm, kg)*10^b*1000

```{r}
# Convert our lengths from mm to cm by dividing by 10
fish<- fish %>% mutate(
  Length_cm= Length_mm/10)

# Column for biomass calculation  W = aL^b
fish<-fish %>% mutate(weight_g = (`Biomass Constant A`*Length_cm^`Biomass Constant B`))

# multiply by number of individuals observed
fish<-fish %>% mutate(Total_Biomass_g = Number*weight_g)

```


# Check and write csv for observation level biomass
```{r}
# check each observation has a biomass value, if all good, write to CSV
fish %>% filter(is.na(Total_Biomass_g)) 

write.csv(fish, file= "../2_outputs/ROV_fish_biomass.csv", row.names = FALSE)

```

# Summary per transect total biomass
```{r}
# Select variables excluding species names etc and Group by transect
biomass.total <- fish %>% select(
Survey,Reef,Site,Transect,T_ID, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity,Total_Biomass_g
)

 
```




```{r}
# Summary per transect total biomass
biomass.total<- biomass.total %>% group_by(Survey,Reef,Site,Transect,T_ID, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity) %>%
  summarise(Total_Biomass_g = sum(Total_Biomass_g))

```

# Write csv for summary total biomass by transect biomass
```{r}

write.csv(biomass.total, file= "../2_outputs/ROV_transect_biomass.csv", row.names = FALSE)

```

