---
title: "05_biomass_calculations"
author: "Gemma Galbraith"
date: "30/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Uses the "ROV_Fish_Final" data to calculate the biomass of each fish observation.

#HouseKeeping
```{R}
rm(list=ls())
#dev.off()
getwd()

```

#load libraries
```{r}
library(tidyverse)
library(rfishbase)
library(lubridate)
```

# Data import
```{r}
fish <-read_csv("../2_outputs/ROV_Fish_Final.csv")

fish <- fish %>% mutate(
  Transect=as.factor(Transect),
  Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Length_mm=as.double(Length_mm),
  Survey=as.factor(Survey),
  Site=as.factor(Site),
  Video = as.factor(Video),
  Comment = as.character(Comment),
  Habitat = as.factor(Habitat),
  Aspect = as.factor(Aspect))

```

# Validate large individuals
Looks like there have been some errors with measuring a bunch of Chromis sp, and other smaller fish. These needs to be filtered out and changed.

The error seems to stem from some observations being entered in cm and some in mm. Therefore, the Length_mm variable is actually a mixture of mm and cm.


```{r}

# Convert our lengths from mm to cm by dividing by 10
fish<- fish %>% mutate(
  Length_cm= Length_mm/10)

# If observed length exceeds max length, assign max length
fish2<-fish %>% mutate(
  Length_cm = case_when(Length_cm > `Max Length (cm)` ~ `Max Length (cm)`,
                        TRUE ~ (Length_cm))
)


# Clear that some lengths are cm and should be x 10. If length_cm is less than 15% of max length x10 
fish2<-fish2 %>% mutate(
  fifteen.per.length  = as.double((`Max Length (cm)`/100)*15),
  Length_cm = case_when(Length_cm < fifteen.per.length ~ (Length_cm)*10,
                        TRUE ~ (Length_cm))
)


# Deal with NAs - this should only apply to Stereo surveys and not to Poms and Chromis species BUT don't apply to anything bigger than 5cm
fish2<-fish2 %>% mutate(
  Length_cm = case_when(is.na(`Max Length (cm)`) & Video == "Stereo" & (grep("Pomacentrus sp|Chromis sp",Species_name, invert = TRUE) & Length_cm <=5)  ~ (Length_cm)*10, TRUE ~ (Length_cm)))

# Random massive labrid - clearly an error

fish2<-fish2 %>% mutate(
  Length_cm = case_when(Length_mm == 1022.5700 & Species_name == "Labridae sp" ~ 10.2,
                        TRUE ~ (Length_cm))
)


```

# Another way to look at what needs fixing

fish3<-fish2 %>% filter(is.na(`Max Length (cm)`)) # the bad stuff that needs fixing
fish4<-fish2 %>% filter(!is.na(`Max Length (cm)`)) # the good stuff

fish3<-fish3 %>% mutate(
  Length_cm = case_when(Video == "Stereo" & (grep("Pomacentrus sp|Chromis sp",Species_name, invert = TRUE)) ~ (Length_cm)*10, TRUE ~ (Length_cm))
)

Then left join back together



# Biomass calcuation for each observation
This takes into account the a and b constants, length in mm and the number of fish/individuals in the observation

W = a x Lb
The units of length and weight in FishBase are centimeter and gram, respectively. Thus when length-weight relationships are not in cm-g, the intercept 'a' is transformed as follows:

a’(cm, g) = a (mm, g)*10^b
a’(cm, g) = a (cm, kg)*1000 
a’(cm, g) = a (mm, mg)*10^b/1000 
a’(cm, g) = a (mm, kg)*10^b*1000

```{r}

# Column for biomass calculation  W = aL^b
fish2<-fish2 %>% mutate(weight_g = (`Biomass Constant A`*Length_cm^`Biomass Constant B`))

# multiply by number of individuals observed
fish2<-fish2 %>% mutate(Total_Biomass_g = Number*weight_g)

```


# Check and write csv for observation level biomass
```{r}
# check each stereo video observation has a biomass value, if all good, write to CSV
fish2 %>% filter(is.na(Total_Biomass_g) & Video == "Stereo") 

fish2<-fish2 %>% select(-Length_mm, - fifteen.per.length)

write.csv(fish2, file= "../2_outputs/ROV_fish_biomass.csv", row.names = FALSE)

```

# Summary per transect total biomass
```{r}
# Select variables excluding species names etc and Group by transect
biomass.total <- fish2 %>% select(
Survey,Reef,Site,Transect,T_ID, Date,Video, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity,Total_Biomass_g
)

 
```



```{r}
# Summary per transect total biomass
biomass.total<- biomass.total %>% group_by(Survey,Reef,Site,Transect,T_ID,Video, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity) %>%
  summarise(Total_Biomass_g = sum(Total_Biomass_g))

```

# Write csv for summary total biomass by transect biomass
```{r}

write.csv(biomass.total, file= "../2_outputs/ROV_transect_biomass.csv", row.names = FALSE)

```

