---
title: "06_biomass_trophic"
author: "GF Galbraith"
date: '2022-06-21'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This script works with the observation level biomass calculated in "05_biomass_calculations". It is used to summarise biomass by trophic or funcational groups.

#HouseKeeping
```{R}
rm(list=ls())
#dev.off()
getwd()
```

#load libraries
```{r}
library(tidyverse)
library(rfishbase)
```

# Data import
```{r}
fish <-read_csv("../2_outputs/ROV_fish_biomass.csv")

fish <- fish %>% mutate(
  Transect=as.factor(Transect),
  Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Length_mm=as.double(Length_mm),
  Survey=as.factor(Survey),
  Site=as.factor(Site),
  Comment = as.character(Comment),
  Date = as.Date(Date),
  Habitat = as.factor(Habitat),
  Aspect = as.factor(Aspect))

```


# Biomass by trophic group - find NA values
```{r}
summary(is.na(fish$`Trophic Group`))
# remove from main data 
fish.2<-fish %>% filter(!is.na(`Trophic Group`))
# create a df of missing species - 277 observations
missing<-fish %>% filter(is.na(`Trophic Group`))
```

```{r}
# Subset observations missing only the species (trophic group applied to genus)
unique(missing$Species_name)
# Note: Gnathodentex aureolineatus and Chromis chrysura have no Trophic group - manual apply
missing$`Trophic Group`[missing$Species_name == "Gnathodentex aureolineatus"] <- "invertivore-mobile"
missing$`Trophic Group`[missing$Species_name == "Chromis chrysura"] <- "planktivore"
unique(missing$Species_name)
```

```{r}
# Missing genera
missing.genera<-missing %>% filter(is.na(Genus))  
missing.species<-missing %>% filter(!is.na(Genus))  
```

```{r}
# Apply a best estimate trophic group based on fishbase information.
unique(missing.genera$Family)
unique(fish$`Trophic Group`)

missing.genera$`Trophic Group`[missing.genera$Family == "Labridae"] <- "omnivore"
missing.genera$`Trophic Group`[missing.genera$Family == "Pomacentridae"] <- "planktivore"
missing.genera$`Trophic Group`[missing.genera$Family == "Clupeidae"] <- "planktivore"
missing.genera$`Trophic Group`[missing.genera$Family == "Carangidae"] <- "piscivore"

# Check all values assigned
missing.genera %>% filter(is.na(`Trophic Group`))  

```

# Missing species, trophic grouop based on genus
```{r}
# Apply a best estimate trophic group based on fishbase information.
unique(missing.species$Genus)
unique(fish$`Trophic Group`)
```

```{r}
sp<-unique(missing.species$Genus)
sp %in% fish$Genus

sp.4<-filter(fish, Genus %in% sp)

sp.4<-sp.4 %>% distinct(Species_name, .keep_all = TRUE) %>% 
  select(Species_name,Genus, `Trophic Group`, `Trophic Level`)

sp.4<-sp.4 %>% filter(!is.na(`Trophic Group`)) 

# Now summerise by the most frequent value of the Trophic group and a mean TL
sp.3<-sp.4 %>% group_by(Genus) %>% 
  summarise(`Trophic Level` = mean(`Trophic Level`),
            `Trophic Group` = names(which.max(table(`Trophic Group`))))

```


# Merge sp.4 with the missing species dfs
```{r}
sp.4<-left_join(missing.species, sp.3, by = "Genus") %>% 
  mutate(`Trophic Group` = coalesce(`Trophic Group.x`,`Trophic Group.y`),
         `Trophic Level` = coalesce(`Trophic Level.x`, `Trophic Level.y`)) %>% 
  select(-c(24,25,29,30))

```

# Check reamining NA Values
```{r}
sp.4 %>% filter(is.na(`Trophic Group`))
# 7 Tricky missing genera still remaining - they must have been the only observations from that genus hence no average TL or most frequent TG.
tricky<-sp.4 %>% filter(is.na(`Trophic Group`))
tricky<-tricky$Genus
tricky %in% mermaid.species$Genus

tricky.2<-filter(mermaid.species, Genus %in% tricky)
tricky.2<-tricky.2 %>% distinct(Species_name, .keep_all = TRUE) %>% 
  select(Species_name,Genus, `Trophic Group`, `Trophic Level`)

# Now summerise by the most frequent value of the Trophic group and a mean TL
tricky.2<-tricky.2 %>% group_by(Genus) %>% 
  summarise(`Trophic Level` = mean(`Trophic Level`),
            `Trophic Group` = names(which.max(table(`Trophic Group`))))

# Merge back
tricky<-sp.4 %>% filter(is.na(`Trophic Group`))
tricky<-left_join(tricky, tricky.2, by = "Genus") %>% 
  mutate(`Trophic Group` = coalesce(`Trophic Group.x`,`Trophic Group.y`),
         `Trophic Level` = coalesce(`Trophic Level.x`, `Trophic Level.y`)) %>% 
  select(-c(27:30))

sp.5<-left_join(sp.4, tricky.2, by = "Genus") %>% 
  mutate(`Trophic Group` = coalesce(`Trophic Group.x`,`Trophic Group.y`),
         `Trophic Level` = coalesce(`Trophic Level.x`, `Trophic Level.y`)) %>% 
    select(-c(27:30))


```


# Final merge
Rbind fish.2, missing.genera, and sp.5.
```{r}

fish.3<- rbind(fish.2,missing.genera,sp.5)

# Final check no NA values.
fish.3 %>% filter(is.na(`Trophic Group`))


```

# Write csv

```{r}

write.csv(fish.3, file= "../2_outputs/ROV_trophic_biomass.csv", row.names = FALSE)

```

```{r}
# Summary per transect total biomass
transect.biomass<- fish.3 %>% group_by(Survey,Reef,Site,Transect,T_ID, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity, `Trophic Group`) %>%
  summarise(Total_Biomass_g = sum(Total_Biomass_g))
```

```{r}

write.csv(transect.biomass, file= "../2_outputs/ROV_trophic_biomass_transect.csv", row.names = FALSE)

```



SCRATCH WORK
-----------------------
# Tricky species
Still no match for 8 genera Try fishbase or mermaid to match? Or Manual extraction from webpage.
```{r}
mermaid.species <-read_csv("../0_data/MERMAID_data/fish-species_GFG.csv")
mermaid.species<-mermaid.species %>% select(c(Genus, Species, Family, Species_name, `Trophic Level`, `Trophic Group`))

```


```{r}
sp.4$`Trophic Group`[sp.4$Genus == "Myliobatis"] <- ""
sp.4$`Trophic Level`[sp.4$Genus == "Myliobatis"] <-
sp.4$`Trophic Group`[sp.4$Genus == "Lethrinus"] <- ""
sp.4$`Trophic Level`[sp.4$Genus == "Lethrinus"] <-
sp.4$`Trophic Group`[sp.4$Genus == "Pseudocheilinus"] <- ""
sp.4$`Trophic Level`[sp.4$Genus == "Pseudocheilinus"] <- 
sp.4$`Trophic Group`[sp.4$Genus == "Pseudochromis"] <- ""
sp.4$`Trophic Level`[sp.4$Genus == "Pseudochromis"] <-   
sp.4$`Trophic Group`[sp.4$Genus == "Pseudocoris"] <- ""
sp.4$`Trophic Level`[sp.4$Genus == "Pseudocoris"] <- 
sp.4$`Trophic Group`[sp.4$Genus == "Chaetodontoplus"] <- ""
sp.4$`Trophic Level`[sp.4$Genus == "Chaetodontoplus"] <- 
sp.4$`Trophic Group`[sp.4$Genus == "Synodus"] <- ""
sp.4$`Trophic Level`[sp.4$Genus == "Synodus"] <- 

  
sp.4 %>% filter(is.na(`Trophic Group`))  
sp.4

```

```{r}

# Manual application of appropriate trophic groups based on unique list of missing species names.
missing.species$`Trophic Group`[missing.species$Genus == "Myripristis"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Meiacanthus"] <- "Halichoeres sp"
missing.species$`Trophic Group`[missing.species$Genus == "Scarus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Chromis"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Pseudanthias"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Thalassoma"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Anampses"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Sargocentron"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Cephalopholis"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Halichoeres"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Myliobatis"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Pomacentrus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Acanthurus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Ctenochaetus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Centropyge"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Lethrinus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Pentapodus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Stegastes"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Cirrhilabrus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Hoplolatilus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Coris"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Pseudocheilinus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Canthigaster"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Pseudodax"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Ptereleotris"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Ptereleotris"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Parupeneus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Rhabdamia"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Caesio"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Pseudochromis"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Pseudocoris"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Chaetodontoplus"] <- ""
missing.species$`Trophic Group`[missing.species$Genus == "Synodus"] <- ""


```



# Summarise and Group by trophic group
```{r}
trophic.biomass<-merged.data %>% group_by(Survey,Reef,Site,Transect,T_ID, Date, Lat, Long, Depth, Temperature, Habitat,Aspect,Benthic_Complexity,`Trophic Group`) %>%
  summarise(Total_Biomass_g = sum(Total_Biomass_g))


```
