---
title: "07_Regional_figures"
author: "Gemma Galbraith"
date: "30/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#HouseKeeping
```{R}
rm(list=ls())
#dev.off()
getwd()

```

#load libraries
```{r}
library(tidyverse)
library(rfishbase)
library(vegan)
library(glmmTMB)
library(sjPlot)
library(DHARMa)
library(car)
library(emmeans)
library(tidyverse)
library(broom.mixed)
library(ggpubr)
library(gridExtra)
library(svglite)
library(lubridate)
library(viridis)  
library(RColorBrewer)
library(unikn)
```

# My colours
#F2C249
#E6772E
#4DB3B3
#E64A45
#3D4C53
#a54a9c
#5074c3

```{r}
my.cols = c("#F2C249", "#E6772E", "#4DB3B3", "#E64A45", "#3D4C53", "#a54a9c", "#5074c3")

seecol(my.cols, 
       col_brd = "black", lwd_brd = 2, 
       title = "See the colors, HEX codes, and RGB values of my.cols") 

depth.cols <- c("#F2C249", "#4DB3B3", "#a54a9c")

seecol(depth.cols, 
       col_brd = "black", lwd_brd = 2, 
       title = "See the colors, HEX codes, and RGB values of my.cols") 
```


# Data import
```{r}
diversity <-read_csv("../0_data/fish_diversity.csv")
biomass <-read_csv("../0_data/fish_biomass.csv")
biomass<-biomass %>% select(T_ID, Total_Biomass_g)

fish<-left_join(diversity, biomass, by="T_ID")

fish <- fish %>% mutate(
  Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Survey=as.factor(Survey),
  Site=as.factor(Site),
  Date = as.Date(Date),
  Habitat = as.factor(Habitat),
  Aspect = as.factor(Aspect),
  Total_biomass_KG = as.double(Total_Biomass_g/1000))

```

# spotted T_ID error
Check BVille transects, some observations do not have a transect number
A fix for now
```{r}
fish <- fish %>% 
  mutate(
    T_ID = case_when(T_ID == "Feb_21_BVille_1_NA" ~ "Feb_21_BVille_1_T1",
                        TRUE ~ as.character(T_ID)),
    T_ID = as.factor(T_ID))
```

# apply Region grouping by reef
```{r}
fish <- fish %>% mutate ( 
  Region = as.factor(case_when(grepl("Lihou", Reef) ~ "Central Coral Sea",
                   grepl("Flinders", Reef) ~ "Central Coral Sea",
                   grepl("Bougainville", Reef) ~ "Northern Coral Sea",
                   grepl("Osprey", Reef) ~ "Northern Coral Sea",
                   grepl("Chillcot", Reef) ~ "Central Coral Sea",
                   grepl("Herald", Reef) ~ "Central Coral Sea",
                   grepl("Holmes", Reef) ~ "Central Coral Sea",
                   grepl("Willis", Reef) ~ "Central Coral Sea",
                   grepl("Frederick", Reef) ~ "Southern Coral Sea",
                   grepl("Kenn", Reef) ~ "Southern Coral Sea",
                   grepl("Wreck", Reef) ~ "Southern Coral Sea",
                   grepl("Saumarez", Reef) ~ "Southern Coral Sea",
                   grepl("Marion", Reef) ~ "Central Coral Sea",)))

unique(fish$Region)
# Should just be three levels

# Ordered factor for Region
fish <-fish %>% mutate(
  Region = factor(Region, levels= c("Southern Coral Sea","Central Coral Sea","Northern Coral Sea"),ordered = TRUE))
```

# Assign Depth grouping
```{r}
fish <- fish %>% mutate(
  Depth_bin = as.factor(case_when(Depth < 30 ~ "Shallow",
                                  Depth > 30 & Depth < 65 ~ "Mid",
                                  TRUE ~ "Deep"))
)

unique(fish$Depth_bin)

# Ordered factor for Region
fish <-fish %>% mutate(
  Depth_bin = factor(Depth_bin, levels= c("Shallow","Mid","Deep"),
                                         ordered = TRUE))
```


# Calcualte mean indices by region and depth

```{r, glmm1 richness}

# Is there a difference between richness across depth bands and regions?
set.seed(123)
glmm1<- glmmTMB(richness ~ Depth_bin+Region+(1|Reef),data = na.omit(fish), family=nbinom1(), REML = TRUE)

simulateResiduals(glmm1, plot=TRUE) # fit is fine

summary(glmm1)
Anova(glmm1) # season is not a sig predictor of fish spawned count

means<-emmeans(glmm1, pairwise~ Depth_bin+Region, transform="response") %>% 
  confint()

means

```

# Richness Figure - boxplot
```{r}
ggplot(fish, aes(y=richness, x=Depth_bin)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)+
  scale_fill_viridis(discrete = TRUE, option = "plasma", direction= -1, alpha =0.7) +
  ylab("Richness (\u00B1 SE)")+
  xlab("Depth")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 50))+ 
  facet_wrap(~Region)+
  theme_classic()+
  theme(axis.text.x = element_text(size=14, color="black"),
        axis.text.y = element_text(size=14, color="black"),
        axis.title.y = element_text(size=16, vjust=2),
        axis.title.x = element_text(size=16, vjust=-1),
        legend.position = "none",
        plot.margin=unit(c(15,5,5,10),"mm"))


```


# Richness Figure - bar
```{r}
plot.means<-as.data.frame(means$emmeans)

tab_df(plot.means)

# Bar
P1<-ggplot(plot.means, aes(y=response, x=Depth_bin)) + 
    geom_bar(aes(fill=Depth_bin),position="stack", stat="identity", width=0.5, alpha = 0.9)+
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), size= 1,width=0.3, alpha = 0.9)+
  scale_fill_manual(values = depth.cols)+
  ylab(expression(Richness~("Species 150m"^-2 %+-%SE)))+  xlab("Depth")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 25))+ 
  facet_wrap(~Region)+
  theme_classic()+
  theme(axis.text.x = element_text(size=14, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=14, color="black"),
        axis.title.y = element_text(size=16, vjust=2),
        axis.title.x = element_text(size=16, vjust=-0.2),
        legend.position = "none",
        strip.text = element_text(size = 14, colour = "black", face = "bold"),
        strip.background = element_blank())


        #plot.margin=unit(c(15,5,5,10),"mm"))



P1

ggsave(filename = "../3_figures/depth_region_richness.png", plot= P1,
       width =200, height=120 , units = "mm", dpi = 500, device = "png" )

```


```{r, glmm2 abundance}

# Is there a difference between abundance across depth bands and regions?
set.seed(123)
glmm2<- glmmTMB(abundance ~ Depth_bin+Region(1|Reef),data = na.omit(fish), family=nbinom1(), REML = TRUE)

simulateResiduals(glmm1, plot=TRUE) # fit is fine

summary(glmm2)
Anova(glmm2) 

means<-emmeans(glmm2, pairwise~ Depth_bin+Region, transform="response") %>% 
  confint()

means

```
# Abundance Figure - bar
```{r}
plot.means<-as.data.frame(means$emmeans)

tab_df(plot.means)

# Bar
 P2<-ggplot(plot.means, aes(y=response, x=Depth_bin)) + 
    geom_bar(aes(fill=Depth_bin),position="stack", stat="identity", width=0.5, alpha = 0.9)+
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), size= 1,width=0.3, alpha = 0.9)+
  scale_fill_manual(values = depth.cols)+
  ylab(expression(Density~("n 150m"^-2 %+-%SE)))+  xlab("Depth")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 250))+ 
  facet_wrap(~Region)+
  theme_classic()+
  theme(axis.text.x = element_text(size=14, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=14, color="black"),
        axis.title.y = element_text(size=16, vjust=2),
        axis.title.x = element_text(size=16, vjust=-0.2),
        legend.position = "none",
        strip.text = element_text(size = 14, colour = "black", face = "bold"),
        strip.background = element_blank())


        #plot.margin=unit(c(15,5,5,10),"mm"))



P2

ggsave(filename = "../3_figures/depth_region_abundance.png", plot= P2,
       width =200, height=120 , units = "mm", dpi = 500, device = "png" )

```

```{r, glmm3 shannon diversity}

# Is there a difference between abundace across depth bands and regions?
set.seed(123)
glmm3<- glmmTMB(shannon ~ Depth_bin+Region+(1|Reef),data = na.omit(fish), family=gaussian(), REML = TRUE)

simulateResiduals(glmm3, plot=TRUE) # fit is fine

summary(glmm3)
Anova(glmm3) 

means<-emmeans(glmm3, pairwise~ Depth_bin+Region, transform="response") %>% 
  confint()

means

```
# Diversity Figure - bar
```{r}
plot.means<-as.data.frame(means$emmeans)

tab_df(plot.means)

# Bar
 P3<-ggplot(plot.means, aes(y=emmean, x=Depth_bin)) + 
    geom_bar(aes(fill=Depth_bin),position="stack", stat="identity", width=0.5, alpha = 0.9)+
  geom_errorbar(aes(ymin=emmean-SE, ymax=emmean+SE), size= 1,width=0.3, alpha = 0.9)+
  scale_fill_manual(values = depth.cols)+
  ylab(expression(Diversity~("H' 150m"^-2 %+-%SE)))+  xlab("Depth")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 4))+ 
  facet_wrap(~Region)+
  theme_classic()+
  theme(axis.text.x = element_text(size=14, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=14, color="black"),
        axis.title.y = element_text(size=16, vjust=2),
        axis.title.x = element_text(size=16, vjust=-0.2),
        legend.position = "none",
        strip.text = element_text(size = 14, colour = "black", face = "bold"),
        strip.background = element_blank())


        #plot.margin=unit(c(15,5,5,10),"mm"))



P3

ggsave(filename = "../3_figures/depth_region_diversity.png", plot= P3,
       width =200, height=120 , units = "mm", dpi = 500, device = "png" )

```

```{r, glmm4 biomass}

# Is there a difference between abundace across depth bands and regions?
set.seed(123)
glmm4<- glmmTMB(Total_biomass_KG ~ Depth_bin+Region+(1|Reef),data = na.omit(fish), family=nbinom1(), REML = TRUE)

simulateResiduals(glmm4, plot=TRUE) # fit is fine

summary(glmm4)
Anova(glmm4) 

means<-emmeans(glmm4, pairwise~ Depth_bin+Region, transform="response") %>% 
  confint()

means

```


# Biomass Figure - bar
```{r}
plot.means<-as.data.frame(means$emmeans)

tab_df(plot.means)

# Bar

 P4<-ggplot(plot.means, aes(y=response, x=Depth_bin)) + 
    geom_bar(aes(fill=Depth_bin),position="stack", stat="identity", width=0.5, alpha = 0.9)+
  geom_errorbar(aes(ymin=response-SE, ymax=response+SE), size= 1,width=0.3, alpha = 0.9)+
  scale_fill_manual(values = depth.cols)+
  #ylab(expression(Biomass~(Kg~150^2~\u00B1~SE)))+
  ylab(expression(Biomass~("Kg 150m"^-2 %+-%SE)))+
  xlab("Depth")+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 60))+ 
  facet_wrap(~Region)+
  theme_classic()+
  theme(axis.text.x = element_text(size=14, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=14, color="black"),
        axis.title.y = element_text(size=16, vjust=2),
        axis.title.x = element_text(size=16, vjust=-0.2),
        legend.position = "none",
        strip.text = element_text(size = 14, colour = "black", face = "bold"),
        strip.background = element_blank())


        #plot.margin=unit(c(15,5,5,10),"mm"))



P4

ggsave(filename = "../3_figures/depth_region_biomass.png", plot= P4,
       width =200, height=120 , units = "mm", dpi = 500, device = "png" )

```


