---
title: "09_BRUV_depth_reef_figures"
output: html_document
date: '2022-07-06'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#HouseKeeping
```{R}
rm(list=ls())
#dev.off()
getwd()

```

#load libraries
```{r}
library(tidyverse)
library(rfishbase)
library(vegan)
library(glmmTMB)
library(sjPlot)
library(DHARMa)
library(car)
library(emmeans)
library(tidyverse)
library(broom.mixed)
library(ggpubr)
library(gridExtra)
library(svglite)
library(lubridate)
library(viridis)  
library(RColorBrewer)
library(unikn)
```

# My colours FOR CSMP ROV/BRUV Depth comparisons - 3 bands
#F0C165 - CREAM CAN
#F3A292 - WEWAK
# 65BAC4 - FOUNTAIN BLUE
#7AAEB2 - GULF STREAM
#4C6274 - BLUE BAYOUX
#95ACBE - NEPAL


```{r, colours}

region.cols <- c("#F3A292","#F0C165", "#65BAC4")
depth.cols <-  c("#95ACBE","#7AAEB2", "#4C6274" )
depth.cols2 <- c( "#7AAEB2","#4C6274")

seecol(depth.cols, 
       col_brd = "black", lwd_brd = 2, 
       title = "See the colors, HEX codes, and RGB values of my.cols") 
```

# Data import
```{r,data}
fish <-read_csv("../2_outputs/BRUV_fish_diversity.csv")
drops <-read_csv("../2_outputs/BRUV_drop_diversity.csv")

fish <- fish %>% mutate(
  Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Survey=as.factor(Survey),
  Site=as.factor(Site),
  Date = as.Date(Date),
  Genus = as.factor(Genus),
  Family = as.factor(Family),
  Species_name = as.factor(Species_name),
  Habitat = as.factor(Habitat),
  Aspect = as.factor(Aspect))

drops<- drops %>% mutate( Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Survey=as.factor(Survey),
  Site=as.factor(Site))

```

# Quick EDA
```{r}
summary(fish)
glimpse(fish)


```

# apply Region grouping by reef
```{r}
levels(fish$Reef)

fish <- fish %>% mutate ( 
  Region = as.factor(case_when(grepl("Lihou", Reef) ~ "Central Coral Sea",
                   grepl("Flinders", Reef) ~ "Central Coral Sea",
                   grepl("Bougainville", Reef) ~ "Northern Coral Sea",
                   grepl("Osprey", Reef) ~ "Northern Coral Sea",
                   grepl("Ashmore", Reef) ~ "Northern Coral Sea",
                   grepl("Chilcott", Reef) ~ "Central Coral Sea",
                   grepl("Herald", Reef) ~ "Central Coral Sea",
                   grepl("Holmes", Reef) ~ "Central Coral Sea",
                   grepl("Willis", Reef) ~ "Central Coral Sea",
                   grepl("Frederick", Reef) ~ "Southern Coral Sea",
                   grepl("Kenn", Reef) ~ "Southern Coral Sea",
                   grepl("Wreck", Reef) ~ "Southern Coral Sea",
                   grepl("Saumarez", Reef) ~ "Southern Coral Sea",
                   grepl("Marion", Reef) ~ "Central Coral Sea",)))

unique(fish$Region)
# Should just be three levels

# Ordered factor for Region
fish <-fish %>% mutate(
  Region = factor(Region, levels= c("Southern Coral Sea","Central Coral Sea","Northern Coral Sea"),ordered = TRUE))
```

# Assign Depth grouping (3 levels)
```{r}
fish <- fish %>% mutate(
  Depth_bin = as.factor(case_when(Depth < 30 ~ "Shallow",
                                  Depth > 30 & Depth < 65 ~ "Mid",
                                  TRUE ~ "Deep"))
)

unique(fish$Depth_bin)

# Ordered factor for depth
fish <-fish %>% mutate(
  Depth_bin = factor(Depth_bin, levels= c("Shallow","Mid","Deep"),
                                         ordered = TRUE))
```

# Assign Depth grouping (2 levels)
```{r}
fish <- fish %>% mutate(
  Depth_bin2 = as.factor(case_when(Depth < 30 ~ "Shallow",
                                   TRUE ~ "Deep")))

unique(fish$Depth_bin2)

# Ordered factor for depth
fish <-fish %>% mutate(
  Depth_bin2 = factor(Depth_bin2, levels= c("Shallow","Deep"),
                                         ordered = TRUE))
```
# Assign Depth grouping (2 levels - different limits)
```{r}
fish <- fish %>% mutate(
  Depth_bin3 = as.factor(case_when(Depth < 40 ~ "Shallow",
                                   TRUE ~ "Deep"))
)

unique(fish$Depth_bin3)

# Ordered factor for depth
fish <-fish %>% mutate(
  Depth_bin3 = factor(Depth_bin3, levels= c("Shallow","Deep"),
                                         ordered = TRUE))
```


# Figure - Richness by Reefs
```{r}
ggplot(fish, aes(y=richness, x=Reef, fill = Region)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)+
  scale_fill_manual(values = region.cols)+
    scale_y_continuous(expand = c(0, 0), limits = c(0, 60))+ 
  ylab("Richness (taxa.drop-1)")+
  xlab("")+
  facet_grid(cols = vars(Region), scales = "free", space ="free",labeller = labeller(Region = label_wrap_gen(width = 10)))+
  theme_classic()+
  theme_classic()+
  theme(axis.text.x = element_text(size=12, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=12, color="black"),
        axis.title.y = element_text(size=14, vjust=2),
        axis.title.x = element_text(size=14, vjust=-1),
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(15,5,5,10),"mm"))
```

# Figure - Abundance by depth and Reefs
```{r}
ggplot(fish, aes(y=abundance, x=Reef, fill = Region)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)+
  scale_fill_manual(values = region.cols)+
   # scale_y_continuous(expand = c(0, 0), limits = c(0, 60))+ 
  ylab("Density (individuals.drop-1)")+
  xlab("")+
   facet_grid(cols = vars(Region), scales = "free", space ="free",labeller = labeller(Region = label_wrap_gen(width = 10)))+
  theme_classic()+
  theme(axis.text.x = element_text(size=12, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=12, color="black"),
        axis.title.y = element_text(size=14, vjust=2),
        axis.title.x = element_text(size=14, vjust=-1),
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(15,5,5,10),"mm"))
```
# Figure - Shannon Diversity by depth and Reefs
```{r}
ggplot(fish, aes(y=shannon, x=Reef, fill = Region)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)+
  scale_fill_manual(values = region.cols)+
  scale_y_continuous(expand = c(0, 0), limits = c(0, 5))+ 
  ylab("Diversity (Shannon H'.drop-1)")+
  xlab("")+
  facet_grid(cols = vars(Region), scales = "free", space ="free",labeller = labeller(Region = label_wrap_gen(width = 10)))+
  theme_classic()+
  theme_classic()+
  theme(axis.text.x = element_text(size=12, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=12, color="black"),
        axis.title.y = element_text(size=14, vjust=2),
        axis.title.x = element_text(size=14, vjust=-1),
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(15,5,5,10),"mm"))
```
# Figure - Richness by Reefs and Depth (2 bands)
```{r}
ggplot(fish, aes(y=richness, x=Depth_bin2, fill = Depth_bin2)) + 
  geom_boxplot(outlier.colour="black", outlier.shape=16,
             outlier.size=2, notch=FALSE)+
  scale_fill_manual(values = depth.cols2)+
    scale_y_continuous(expand = c(0, 0), limits = c(0, 60))+ 
  ylab("Richness (taxa.drop-1)")+
  xlab("")+
  facet_grid(cols = vars(Region, Reef), scales = "free", space ="free",labeller = labeller(Region = label_wrap_gen(width = 10)))+
  theme_classic()+
  theme_classic()+
  theme(axis.text.x = element_text(size=12, color="black", angle = 45, vjust = 0.6),
        axis.text.y = element_text(size=12, color="black"),
        axis.title.y = element_text(size=14, vjust=2),
        axis.title.x = element_text(size=14, vjust=-1),
        legend.position = "none",
        strip.background = element_blank(),
        plot.margin=unit(c(15,5,5,10),"mm"))
```


    