---
title: "ROV_fish_wrangle"
author: "Gemma Galbraith"
date: "26/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#HouseKeeping
```{R}
rm(list=ls())
#dev.off()
#getwd()

```

#load libraries
```{r}
library(tidyverse)
library(rfishbase)
```

#Import fish data
```{r}
fish <-read_csv("../0_data/ROV_fish_data.csv")

fish <- fish %>% mutate(
  Transect=as.factor(Transect),
  Reef=as.factor(Reef),
  T_ID=as.factor(T_ID),
  Length_mm=as.double(Length_mm),
  Survey=as.factor(Survey),
  Site=as.factor(Site))

```

#Transect infomation csv.
Take the output file "transect_summary_data" and merge by TID. This now assigns the details e.g. depth, temp, complexity etc to each observation

```{r}
transect_data <-read_csv("../0_data/Feb21_Transect_Summary.csv")

transect_data<-select(transect_data,c(-Day_no, -Dive_no, -Trash, -Cucumbers,- Notes))

transect_data$Transect<-sub("^", "T", transect_data$Transect)

# Redefine some variables
transect_data <- transect_data %>% mutate(
  Reef=as.factor(Reef),
  Site=as.factor(Site),
  Transect=as.factor(Transect),
  Date=as.Date(Date, format = "%d/%m/%Y"),
  Habitat=as.factor(Habitat),
  Aspect=as.factor(Aspect),
  Survey=as.factor("Feb_21"),
  Site=as.factor(Site),
  T_ID = paste(Survey,Site,Transect, sep = '_'),
  T_ID=as.factor(T_ID))
```

#Merge fish and transect data - prechecks for matching transects
```{r}

#  check you have correct merge by unique values for T_ID - this is critical
distinct(transect_data, T_ID) # 135 transects
distinct(fish, T_ID) #128 rows - some transects had no fish so this is fine
# Check fish observations that don't have a transect number
which(is.na(fish$Transect))

fish %>% filter(is.na(Transect))
fish %>% filter(is.na(T_ID))

# There were some NA values - I had to manually go back through the csv files and videos to check.

levels(fish$T_ID)
unique(fish$T_ID) #128 unique transect IDs in the fish
unique(transect_data$T_ID) #135 transects in the transect_data
# This means we have 7 transects analysed which had no fish, could be lagoons ect
# Needs to be checked, do this using an anti-join - will select all rows from df1 (transect_data) that are not present in df2 (fish)

transect_data %>% anti_join(fish, by = "T_ID")
fish %>% anti_join(transect_data, by = "T_ID")
levels(transect_data$T_ID)
# 26/05/22 Saumarez, Wreck, Kenn, Marion and Frederick data are not yet in the project. Others in lagoon habitat are likely true zero observation transects. This should be made note of in the transect summary as the video anlaysis is done.
# Discrepancy between fish Site BVille and transect Site Bville. Changed in Transect summary data csv.
# Chilcott was spelt incorrectly inthe Eventmeasure file but correclty in the transect_data file, consequently wasn't matching - changed mannually in csv file now.
```

#The Merge
```{r}

# data <- merge(fish, transect_data, by =c("T_ID"))
# Check that you retain all observations from fish in the merge. I.e "data" should have the same no of obs as "fish" just with the info from "transect_data" appended.

data<- fish %>% 
  left_join(transect_data,by ="T_ID") %>% 
  select(-Reef.y, -Site.y, -Transect.y, -Survey.y) %>% 
  rename(Reef = Reef.x, Survey = Survey.x, Site = Site.x, Transect = Transect.x)

# Removed and rename duplicate variables created in the merge

```

# Missing species

If species names have been added in the "comments" variable in EventMeasure, these need to be pulled into the family, genus , species variables. Some comments have average length estimates for that species appended to the species name. Using regex and or coalesce we can split anything numeric into another variable and eventually transfer to the length_mm variable.

```{r}
# Inspect the comments variable and identify values 
data %>% filter(
  !is.na(Comment)
)

# Remove any observation with a comment value from the main data and create a sep df. The just work on data2.
data2<-data %>% filter(
  !is.na(Comment)
)

data<-data %>% filter(
  is.na(Comment)
)

# Most follow the "species_length.est" pattern but there are some funky entries. Split data2 again.
# Fish species and species with average lengths follow the pattern "species_length". Use this in regex to select
# /[a-z]+\_[0-9]+\.\d|[a-z]+\_[0-9]+/mg - worked in 101 but not here...
# funky_data <- data2 %>% filter(!str_detect(Comment, pattern = "/^\\[a-z]+\\_[0-9]+\\.\\d$|^\\[a-z]+\\_[0-9]+$"))

funky_data <- data2 %>% filter(!str_detect(Comment, pattern = "^[a-z]+\\_[0-9]+|^[0-9]+"))
# funky data is everything else. #

data2 <- data2 %>% filter(str_detect(Comment, pattern = "^[a-z]+\\_[0-9]+|^[0-9]+"))
# data2 now has all the "species_length.est" observations 

single.species <- funky_data %>% filter (str_detect(Comment, pattern = "^[a-z]+$"))
funky_data <- funky_data %>% filter (!str_detect(Comment, pattern = "^[a-z]+$"))

```

# data2 - deal with the "species_length.est" value in comments
```{r}
# First split the comment by the "_"
data2<- data2 %>% separate(
  Comment, c("temp.species","temp.length"),sep = "_", remove = FALSE
) %>% mutate(
  temp.length=as.double(temp.length)
)

data2$Length_mm[116] <- 200
data2$temp.species[116] <- NA

# COALESCE temp.species and temp.length to fill NA values
data2 <- data2 %>% mutate(
  Species=coalesce(Species, temp.species),
  Length_mm=coalesce(Length_mm, temp.length))

data2<- select(data2, c(-temp.species, -temp.length))
```


# Single species names
```{r}
# some comments are just single species names that need to be moved to the species column.
# Then ready to R bind with main data
single.species <- single.species %>% mutate(
  Species=coalesce(Species, Comment))
```

# Dealing with the funky data

This will be slightly different each time you add new survey data to the script. You will need to work through case by case.
https://stackoverflow.com/questions/70868379/replace-one-column-with-another-using-regex-matching-in-r

```{r}

funky_data$Genus[6] <- "Paracentropyge"
funky_data$Species[6] <- "multifasciata"
funky_data$Length_mm[6] <- 10.3

```

# Row bind corrected single species and funky data back to main data
```{r}
fin.data<-rbind(data,data2,single.species,funky_data)
# Check that your final data here has the same observations as "fish" df that you started with. Otherwise you have lost some in the cleaning processes above :?

write.csv(fin.data, file= "../0_data/Final_fish.csv", row.names = FALSE)

```


# Match with MERMAID database for LW constants
```{r}

mermaid.species <-read_csv("../0_data/MERMAID_data/fish-species.csv")

mermaid.species<-mermaid.species %>% mutate(
  Species_name = paste(Genus,Species, sep = " ")
)

fin.data<-fin.data %>% mutate(
  Species_name = paste(Genus,Species, sep = " ")
)

fin.data %>% anti_join(mermaid.species, by = "Species_name")

```


#Fix all the issues! Working off the mERMAID data base
```{r}

# All NA species get "sp"
fin.data$Species[is.na(fin.data$Species)]<- "sp"
# spp <- sp in Species
fin.data$Species[fin.data$Species == "spp"] <- "sp"

# Oxycheilinus diagrammus <- digramma. This needs a final catch all argument for case_when to say ' if it doesn't fit these rules just give it what it had before'.

fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Oxycheilinus" & Species == "digrammus" ~ "digramma",
                        TRUE ~ as.character(Species))
  )

# Paracentropyge multifasciata - added manually to Eventmeasure species files and MERMAID.csv
# Max length 12.0 , a = 0.03236, b = 2.89, trophic level = 2.8

# Chaetodon mertensi - added mannually to MERMAID.csv for better future use
# Chaetodon mertensii - spelling corrected
# martensii - spelling corrected
# meretensii - spelling corrected
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Chaetodon" & Species == "mertensi" ~ "mertensii",
                        TRUE ~ as.character(Species))
  )

fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Chaetodon" & Species == "martensi" ~ "mertensii",
                        TRUE ~ as.character(Species))
  )

fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Chaetodon" & Species == "meretensii" ~ "mertensii",
                        TRUE ~ as.character(Species))
  )

fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Chaetodon" & Species == "martensii" ~ "mertensii",
                        TRUE ~ as.character(Species))
  )

# Gymnocranius eunanus <- euanus - spelling
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Gymnocranius" & Species == "eunanus" ~ "euanus",
                        TRUE ~ as.character(Species))
  )

# Anampses feminius - spelling of femininus
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Anampses" & Species == "feminius" ~ "femininus",
                        TRUE ~ as.character(Species))
  )

# added manually to MERMAID and EM species list
# Max length 24.0 , a = 0.00977, b = 3.07, Trophic level = 3.5

# Genicanthus watanabie - added manually to MERMAID and EM species list
# Max length 15.0 , a = 0.03236, b = 2.89, Trophic level = 3.4

# Acanthurus nubilis  - spelling <- nubilus
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Acanthurus" & Species == "nubilis" ~ "nubilus",
                        TRUE ~ as.character(Species))
  )

# Diagramma pictum/pictum labiosum  - just pictum on MERMAID
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Diagramma" & Species == "pictum labiosum" ~ "pictum",
                        TRUE ~ as.character(Species))
  )

# Chrysiptera starcki -  - added manually to MERMAID and EM species list
# Max length 7.0 , a = 0.01479, b = 2.98, Trophic level = 2.8
#length_weight("Chrysiptera starcki")
# and a spelling mistake - starki
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Chrysiptera" & Species == "starki" ~ "starcki",
                        TRUE ~ as.character(Species))
  )


# Pseudanthias ventralis
# Max length 7.0 , a = 0.01259, b = 3.01, Trophic level = 3.3

# Pseudochromis flavipinnis
# Max length 7.2 , a = 0.01148, b = 3.01, Trophic level = 3.5

# Sufflamen chrysopterus - spelling <- chrysopterum
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Sufflamen" & Species == "chrysopterus" ~ "chrysopterum",
                        TRUE ~ as.character(Species))
  )

# Chaetodon pelwensis - spelling <-pelewensis
# Max length 12.5 , a = 0.02089, b = 3.02, Trophic level = 2.9
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Chaetodon" & Species == "pelwensis" ~ "pelewensis",
                        TRUE ~ as.character(Species))
  )

# Pseudocaranx georgianus
# Max length 122 cm , a = 0.01413, b = 2.96, Trophic level = 3.9

# Chromis xanthurus - spelling <- xanthura
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Chromis" & Species == "xanthurus" ~ "xanthura",
                        TRUE ~ as.character(Species))
  )

# Ctenochaetus nigricans - error. Video checked - should be acanthurus nigricans
fin.data <- fin.data %>% 
  mutate(
    Genus = case_when(Genus == "Ctenochaetus" & Species == "nigricans" ~ "Acanthurus",
                        TRUE ~ as.character(Genus))
  )

```

# Pre merge second check

```{r}
# Refresh species_name after changes
fin.data<-fin.data %>% mutate(
  Species_name = paste(Genus,Species, sep = " ")
)


fin.data %>% anti_join(mermaid.species, by = "Species_name")


```


# Same species match wrangle but using the rfishbase list

```{r}

fishbase.data <-read_csv("../0_data/fish_list.csv")
fishbase.data<- fishbase.data %>% rename(
  Species_name = Species
)

fin.data %>% anti_join(fishbase.data, by = "Species_name")

# Pseudocaranx georgianus is not accepted name <- Pseudocaranx dentex
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Pseudocaranx" & Species == "georgianus" ~ "dentex",
                        TRUE ~ as.character(Species))
  )
# Zebrasonma veliferum is not accepted name <- Zebrasoma velifer
fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Zebrasoma" & Species == "veliferum" ~ "velifer",
                        TRUE ~ as.character(Species))
  )

# Spelling issues with Genicanthus watanabei

fin.data %>% filter(
  str_detect(Species, pattern = "wata")
  )

fin.data <- fin.data %>% 
  mutate(
    Species = case_when(Genus == "Genicanthus" & str_detect(Species, "wata") ~ "watanabei",
                        TRUE ~ as.character(Species))
  )

# Pseudochromis flavipinnis < Chrysiptera flavipinnis
fin.data <- fin.data %>% 
  mutate(
    Genus = case_when(Genus == "Pseudochromis" & Species == "flavipinnis" ~ "Chrysiptera",
                        TRUE ~ as.character(Genus))
  )


```

```{r}
# Refresh species_name after changes
fin.data<-fin.data %>% mutate(
  Species_name = paste(Genus,Species, sep = " ")
)


fin.data %>% anti_join(fishbase.data, by = "Species_name")


```

# Merge with fishbase
```{r}



```



